Adaption of the Chumbley Score to matching of bullet striation marks

### 2018-03-03 doc/paper.Rmd
introduction, error rates

## Motivation
Same source analyses are a major part of an Forensic Toolmark Examiner's job. In current practice examiners  make these comparisons visually under a comparison microscope and come to one of the following four conclusions: identification, inconclusive, elimination or unsuitable for examination \citet{afte-website-2}. These conclusions are made on the basis of "unique surface contours" of the two toolmarks being in "sufficient agreement" \citep{afte-toolmarks1998}. AFTE describes the term "sufficient agreement" as the possibility of another tool producing the markings under comparison, as practically impossible \citep{afte-toolmarks1998} \citep{afte-website}. 
%\gk{Although, they go on to concede that the interpretation of the identification depends on the examiner and is  therefore subjective in nature.} 
This subjectivity in the assessment as well as the lack of error rates are the main points of criticisms in the toolmark examination first raised by the National Research Council in 2009 \citep{NAS:2009} and later emphasized further by the President's Council of Advisors on Science and Technology \citep{pcast2016}.

@@

\citet{manytoolmarks2} define a relative distance metric and use it as similarity measure between two toolmarks. \citet{manytoolmarks1} extract many small segments in the markings of two toolmarks and compare similiarity using a maximum pearson correlation coefficient. 
The Chumbley scoring method, first introduced by \citet{chumbley}, uses a similar but more extensive framework based on a Mann-Whitney U test of the resulting correlation coefficients. This approach was later improved by Hadler and Morris \citep{hadler}. \hh{XXX describe the improvement in half a sentence XXX} In this paper, we are investigating the applicability of the Chumbley scoring method by \citet{hadler} to assess striation marks on bullet lands for same-source identification.

Striation marks on bullets are made by impurities in the barrel. As the bullet travels through the barrel, these imperfections leave "scratches" on the bullet surface. Typically, only striation marks in the land engraved areas (LEAs) are considered \citet{afte-article1992}. Bullet lands are depressed areas between the  grooves made by the rifling action of the barrel. Compared to toolmarks made by screwdrivers striation marks on bullets are typically much smaller, both in length and in width. Bullets also have a curved cross-sectional topography \hh{sketch?}.

Bullet matching methods depend on the attribute being used for making comparisons. For single feature based matching \citet{chu2013} used an automatic method for counting consequtive matching striae (CMS). They reported zero false positive rate but do not discuss about the false negative rates for land to land matches. \citet{ma2004} and \citet{vorburger2011} discuss about CCF (cross-correlation function) and its discriminating power and application but do not talk about matches and non-matches. \citet{aoas} use multiple attributes like CCF, CMS, D (distance measure) etc in a random forrest based method and compare every land against every other land of the Hamby dataset \citet{hamby}. They report an out-of-bag error rate with  false positive error rate of 0.0001 and false negative error rate of 0.2267 for 100 trees. For 300 trees they report correct prediction of all matches and non-matches.

The Chumbley score  provides us with another approach in the same-source assessment of bullet striation marks.
\citet{chumbley}, in their paper, compare two toolmarks with the intention of determining if it comes from the same tool. The data for their study was obtained from 50 sequentially manufactured screwdriver tips.\citet{chumbley} reported error rates for marking made by the tips at differnt angles. For a 30 degree they showed false negative error rate to be 0.023 and  false positive error rate be 0.09. For other angles the error rates remain the same for false negatives and 0.01 for false positives. The improvement proposed by \citet{hadler} on the other hand, did not consider effect of angles and gave 0.06 false  negative error rate and 0 false poisitive error rate for markings made at the same angle.

\gk{\citep{afte-website} quotes sufficient agreement as "agreement of individual characteristics is of a quantity and quality that the likelihood another tool could have made the mark is so remote as to be considered a practical impossibility"} 
\gk{could not find the pdf for AFTE theory of identification as it relates to toolmarks pdf. I believe it is not available for free. But from the AFTE Journal the abstract was available which mentions the term " sufficient agreement" as quoted above. Currently the website seems like the most reliable way of quoting AFTE, i have checked other sources in both occasions where the website has been quoated above, like an FBI paper available at the FBI website \citet{fbi-paper} and some non published but available pdf files in the AFTE website that have used these lines and quoted the AFTE journal. But since I cannot verify it in the Journal I have quoated the website.}
\gk{faden et al have not cited any paper that discusses the algorithm, i assume it was developed for the paper, the discussion of which is not as extensive as chumbley's. Its a variation of the chumbley method but from my understanding it is just the optimization step. and this paper came out much before the chumbley score.}

<!-- #### -->

\hh{XXX first sentence: Same source analyses are a major part of an Forensic Toolmark examiners job. XXX second sentence: describe standard practice: In current practice firearm examiners  make these comparisons visually under a comparison microscope and report XXX what is being reported: AFTE. XXX find: AFTE theory of identification as it relates to toolmarks  }
\hh{One of the main criticisms in this toolmark examination first raised by the National Research Council in 2009 [@NAS:2009] and later emphasized further by the President's Council of Advisors on Science and Technology [@pcast2016] are the amount of subjectivity in the assessment and the lack of error rates. }
\hh{XXX we have digitized images, e.g. profilometers or Confocal microscopy allow us to measure the surface [@vorburger2016]}

\hh{Using a statistical approach removes both subjectivity from the assessment and allows us to provide error rates for both false positive and false negative identifications. }
\hh{Multiple toolmarks have been studied in the literature:  \citep{manytoolmarks1, manytoolmarks2, chumbley} have studied screwdriver marks digitized using a profilometer,  \citep{manytoolmarks2} investigate groove pliers XXX data? XXX  to slip-joint pliers \citep{afte-chumbley} XXX. In all of these cases, the statistical test used, is a variation of the Chumbley scoring method, first introduced by XXX and later improved by XXX. }

\hh{XXX In this paper, we are investigating the applicability of the Chumbley scoring method by @hadler to assessing striation marks on bullets for same-source identification. }

\hh{Striation marks on bullets are made by impurities in the barrel. As the bullet travels through the barrel, these imperfections leave "scratches" on the bullet surface. Typically, only striation marks in the land engraved areas (LEAs) are considered (XXX citation) [@afte-article1992]. XXX explain lands and grooves XXX. Compared to toolmarks made by screwdrivers striation marks on bullets are typically much smaller, both in length and in width. Bullets also have a curved cross-sectional topography. 
}

\hh{XXX How are bullets matched now? \citet{chu2013}: CMS, automatically - error rate?, using ccf XXX Vorburger paper XXX  and \citet{aoas}: multiple features, error rates }

\hh{The Chumbley score  provides us with another approach in the same-source assessment of bullet striation marks. }

>>>>

## Motivation
Same source analyses are a major part of an Forensic Toolmark Examiner's job. In current practice examiners  make these comparisons visually under a comparison microscope and come to one of the following four conclusions: identification, inconclusive, elimination or unsuitable for examination \citep{afte-toolmarks1998}. These conclusions are made on the basis of "unique surface contours" of the two toolmarks being in "sufficient agreement" \citep{afte-toolmarks1998}. AFTE describes the term "sufficient agreement" as the possibility of another tool producing the markings under comparison, as practically impossible \citep{afte-toolmarks1998}. This subjectivity in the assessment as well as the lack of error rates are the main points of criticisms in the toolmark examination first raised by the National Research Council in 2009 \citep{NAS:2009} and later emphasized further by the President's Council of Advisors on Science and Technology \citep{pcast2016}.

@@

\citet{manytoolmarks2} define a relative distance metric and use it as similarity measure between two toolmarks. \citet{manytoolmarks1} extract many small segments in the markings of two toolmarks and compare similiarity using a maximum pearson correlation coefficient. 
The Chumbley scoring method, first introduced by \citet{chumbley}, uses a similar but more extensive framework based on a Mann-Whitney U test of the resulting correlation coefficients. This approach was later improved by Hadler and Morris \citep{hadler} where they introduce a deterministic approach and address the issue of lack of independece between segements of striae.  In this paper, we are investigating the applicability of the Chumbley scoring method by \citet{hadler} to assess striation marks on bullet lands for same-source identification.

Striation marks on bullets are made by impurities in the barrel. As the bullet travels through the barrel, these imperfections leave "scratches" on the bullet surface. Typically, only striation marks in the land engraved areas (LEAs) are considered \citet{afte-article1992}. Bullet lands are depressed areas between the  grooves made by the rifling action of the barrel. Compared to toolmarks made by screwdrivers striation marks on bullets are typically much smaller, both in length and in width. Bullets also have a curved cross-sectional topography \hh{sketch?}. Figure \ref{fig:rgl} shows us how the signature from a bullet land lines up with the image of the land from which it was extracted. We can see also see in the figure how the depth and relative position of the striation markings seen in the image are interpreted as the signature.

Bullet matching methods depend on the attribute being used for making comparisons. For single feature based matching \citet{chu2013} used an automatic method for counting consequtive matching striae (CMS). They reported an error rate 52% of the known same source lands comparisons as misidentified and zero false positives for known different source lands. \citet{ma2004} and \citet{vorburger2011} discuss about CCF (cross-correlation function) and its discriminating power and application but do not talk about matches and non-matches. \citet{aoas} use multiple attributes like CCF, CMS, D (distance measure) etc in a random forrest based method and compare every land against every other land of the Hamby dataset \citet{hamby}. They report an out-of-bag error rate of 0.0046.

The Chumbley score  provides us with another approach in the same-source assessment of bullet striation marks.\citet{chumbley}, in their paper, compare two toolmarks with the intention of determining if it comes from the same tool. The data for their study was obtained from 50 sequentially manufactured screwdriver tips.\citet{chumbley} reported error rates for marking made by the tips at differnt angles. For a 30 degree they report false negative error rate to be 0.023 and  false positive error rate be 0.09. For other angles the error rates remain the same for false negatives and 0.01 for false positives. The improvement proposed by \citet{hadler} on the other hand, did not consider effect of angles and gave 0.06 false  negative error rate and 0 false poisitive error rate for markings made at the same angle.
